---
title: "Scraping butte co. sheriff website data"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
# create image folder ----
if (!file.exists("images/")) {
    dir.create("images/")
}
# create data folder ----
if (!file.exists("data/")) {
    dir.create("data/")
}
# create meta folder ----
if (!file.exists("meta/")) {
    dir.create("meta/")
}
# knitr settings ------
knitr::opts_chunk$set(
    echo = TRUE, # show all code
    tidy = FALSE, # cleaner code printing
    size = "small",
    fig.path = "images/") # smaller code
knitr::opts_knit$set(
    width = 78)
base::options(tibble.print_max = 25,
              tibble.width = 78)
```

# Motivation

Collect crime data in Butte county after 2018 (i.e. after the wildfires on November 8th)

## Website

Using the selector gadget, we scraped the html page found here:

https://bcso.crimegraphics.com/2013/default.aspx?InitialTab=10

```{r packages}
library(rvest)
library(httr)
library(xml2)
```

## Warning 

> "This website is provided as a public service by the Butte County Sheriff's Office. In an effort to protect victim privacy this website does not provide information regarding juvenile offenders, or specific information regarding calls for service that are sensitive in nature. The Department does not guarantee the accuracy, completeness, or timeliness of the information contained on this website regarding specific incidents, crimes, or people with respect to the omission of information that may have not yet been filed or is pending filing with a court of jurisdiction relating to criminal offenses."

## Extract data

This code will extract the data from the html.

```{r download-html}
sheriff_url <- "https://bcso.crimegraphics.com/2013/default.aspx?InitialTab=10"
# read html
sheriff_html <- xml2::read_html("https://bcso.crimegraphics.com/2013/default.aspx?InitialTab=10")
str(sheriff_html)
str(sheriff_html$node)
str(sheriff_html$doc)
```

Convert to node.

```{r sheriff_vector}
sheriff_vector <- sheriff_html %>% 
    rvest::html_nodes("#Bull") %>% 
    html_text() 
# sheriff_vector %>% head()
```

I use the `stringr::str_view_all` to find the pattern, but now I want to split this into a data frame. 


```{r locate-string-pattern}
stringr::str_view_all(sheriff_vector,
                          # match number format of 00:00
                          pattern = "\\d\\d:\\d\\d")
```

Now split at this point.

This was sent to RStudio Community:

https://community.rstudio.com/t/split-character-vector-into-tibble-using-regex-but-remove-nothing/28282

And two options for dealing with this problem. I am using the option found [here](https://www.r-bloggers.com/strsplit-but-keeping-the-delimiter/).

```{r strsplit_keep}
strsplit_keep <- function(x,
                     split,
                     type = "remove",
                     perl = FALSE,
                     ...) {
  if (type == "remove") {
    # use base::strsplit
    out <- base::strsplit(x = x, split = split, perl = perl, ...)
  } else if (type == "before") {
    # split before the delimiter and keep it
    out <- base::strsplit(x = x,
                          split = paste0("(?<=.)(?=", split, ")"),
                          perl = TRUE,
                          ...)
  } else if (type == "after") {
    # split after the delimiter and keep it
    out <- base::strsplit(x = x,
                          split = paste0("(?<=", split, ")"),
                          perl = TRUE,
                          ...)
  } else {
    # wrong type input
    stop("type must be remove, after or before!")
  }
  return(out)
}
```


See if a simple look behind and keep works. 

```{r}
regex_pattern <- paste0("(?<=.)(?=", "(\\d\\d:\\d\\d)", ")")
regex_pattern
```


```{r ButteLogRaw, message=FALSE, warning=FALSE}
ButteLogRaw <- strsplit_keep(x = sheriff_vector, 
              split = "(\\d\\d:\\d\\d)", 
              type = "before") %>% 
    as_tibble(.name_repair = "unique") %>% 
    dplyr::rename(time = `...1`)
```

Now I can wrangle this into shape with two variables, `call_time` and `description`.

```{r ButteLog-call_time-description, message=FALSE, warning=FALSE}
ButteLog <- ButteLogRaw %>% 
    tidyr::separate(col = time, 
                    into = c("call_time", "dummy"), 
                    sep = " ",
                    remove = FALSE) %>% 
    dplyr::mutate(description = stringr::str_remove_all(string = time,
                                        pattern = "(\\d\\d:\\d\\d)"))
ButteLog <- ButteLog %>% 
    select(-dummy,
           original_record = time)
ButteLog %>% head()
```

## Separate capitalized text

But I also want to separate the capitalized text from the other text. The regular expression for 

```{r extract}
df <- data.frame(x = c(NA, "a-b", "a-d", "b-c", "d-e"))
df
df %>% tidyr::extract(col = x, into = "A")
df %>% tidyr::extract(col = x, into = c("A", "B"), 
                      regex = "([[:alnum:]]+)-([[:alnum:]]+)")
# If no match, NA:
df %>% tidyr::extract(col = x, 
                      into = c("A", "B"), 
                      regex = "([a-d]+)-([a-d]+)")
```


```{r}
ButteLog %>% 
    dplyr::mutate(activity = 
                      stringr::str_match_all(string = original_record,
                                                      pattern = "[A-Z]"))
```


Export these data.


```{r export-data, eval=TRUE}
# raw
write_csv(as.data.frame(ButteLogRaw), path = paste0("data/raw/",
                         "ButteLogRaw", 
                         base::noquote(lubridate::today()),
                         ".csv"))

# cleaner
write_csv(as.data.frame(ButteLog), path = paste0("data/",
                         "ButteLog", 
                         base::noquote(lubridate::today()),
                         ".csv"))
```


```{r}
writeLines(fs::dir_ls("data/raw/"))
writeLines(fs::dir_ls("data/"))
```

