---
title: "Scraping MMA"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(magrittr)
# create image folder ----
if (!file.exists("images/")) {
    dir.create("images/")
}
# create data folder ----
if (!file.exists("data/")) {
    dir.create("data/")
}
# create meta folder ----
if (!file.exists("meta/")) {
    dir.create("meta/")
}
# knitr settings ------
knitr::opts_chunk$set(
    echo = TRUE, # show all code
    tidy = FALSE, # cleaner code printing
    size = "small",
    fig.path = "images/") # smaller code
knitr::opts_knit$set(
    width = 78)
base::options(tibble.print_max = 25,
              tibble.width = 78)
```

# Motivation

I wanted to build a database of MMA fighters following a method similar to [this post](http://www.fightprior.com/2016/04/29/scrapingMMA/). 

## Scrape a single fighter

Before build a function that could autmate the web scraping, I wanted to scrape a single fighter's data and see how messy it was. The code below scrapes the data for Jon Jones from www.sherdog.com and puts it into data frames. 

```{r fighter_page}
# Load packages
library(rvest)
library(dplyr)
library(xml2)

# read the webpage page of the fighter that we are interested in
jjones_page <- xml2::read_html("https://www.sherdog.com/fighter/Jon-Jones-27944")
jjones_page %>%
  # use CSS selector to extract relevant entries from html
  rvest::html_nodes(".nickname em , .fn") %>%
  # turn the html output into simple text fields
  html_text
```

## Extracting fight history and opponent links

Create `JonesTables` by using the `rvest::html_nodes()` function to find the `css = "table"`. 

```{r JonesTables}
JonesTables <- jjones_page %>%
  # use CSS selector to extract relevant entries from html
  rvest::html_nodes(css = "table")
str(JonesTables)
```

## Fighter info

The information on each fighter is in the 

""

```{r JonesInfo}
# devtools::install_github('dantonnoriega/xmltools')
library(xmltools)
JonesInfo <- jjones_page %>%
  # use CSS selector to extract relevant entries from html
  rvest::html_nodes(css = ".vcard , .count_history , .bio_graph , .birthplace , .size_info , .height , .birthday")
JonesInfo %>% str() 
```



## Fight results

Check the `JonesTables` for the fight results (stored in position `[[1]]`), 

```{r locate-table}
rvest::html_table(JonesTables[[2]], fill = TRUE) %>% 
    utils::head(10)
```


Now we can use `magrittr::set_names()` to assign column names, then filter out row 1 (they are the old column names) and store in `JonesRecords`.

```{r JonesRecords}
# locate tables
JonesRecordsRaw <- rvest::html_table(JonesTables[[2]], fill = TRUE) %>% 
    # col names
    magrittr::set_names(., value = c("result", "fighter", "event",
                                     "method_referee", "rounds", 
                                     "time")) %>% 
    dplyr::filter(result != "Result")
    # remove first row of values 
    
JonesRecordsRaw %>% dplyr::glimpse(78)
```

Now I want to separate out the variables that are combined into a single column. 

```{r JonesRecordsRaw}
JonesRecordsRaw %>% head()
```

I need to extract the date from the `event` column, which I can do with a neat trick using `stringr::str_sub`.

This is a small example of the `event` column from above. I want to see how many characters there are, and then remove the characters that correspond to the date. 

```{r example_event}
example_event <- "UFC 235 - Jones vs. SmithMar / 02 / 2019"
# nchar(example_event) # 40
stringr::str_sub(example_event, -15)
```

```{r events}
JonesRecords <- JonesRecordsRaw %>% 
    dplyr::mutate(event_date = stringr::str_sub(event, -15),
                  event_name = stringr::str_sub(event, end = -16)) %>% 
    dplyr::select(dplyr::contains("event"), 
                  dplyr::everything())
JonesRecords %>% utils::head()
```

## Separate method and referee

The `method` for each fight result is stored in the same column with the `referee`, so I will split these up with `tidyr::separate()`. Then I will remove some extra text and format the date in `event_date`.

```{r method_referee-dates}
# test 
JonesRecords %>% 
    tidyr::separate(col = method_referee,
                    into = c("method", "referee"),
                    sep = "\\)",
                    remove = FALSE) %>% 
    dplyr::mutate(method = stringr::str_replace_all(method, 
                                            pattern = "\\(",
                                            replacement = "= "),
                  event_date = lubridate::mdy(event_date)) %>% 
    dplyr::select(event,
                  event_name,
                  method,
                  event,
                  event_date)
# assign
JonesRecords <- JonesRecords %>% 
    tidyr::separate(col = method_referee,
                    into = c("method", "referee"),
                    sep = "\\)",
                    remove = FALSE) %>% 
    dplyr::mutate(method = stringr::str_replace_all(method, 
                                            pattern = "\\(",
                                            replacement = "= "),
                  event_date = lubridate::mdy(event_date)) %>% 
    # remove old cols
                  dplyr::select(-event,
                                -method_referee)
```

```{r check-JonesRecords}
JonesRecords %>% glimpse(78)
```

Now that I have a time-stamped data set, I will export it as a .csv file. 

```{r export_raw, eval=TRUE}
# raw
write_csv(as.data.frame(JonesRecordsRaw), path = paste0("data/raw/",
                         "JonesRecordsRaw", 
                         base::noquote(lubridate::today()),
                         ".csv"))
writeLines(fs::dir_ls("data/raw/"))

# cleaner
write_csv(as.data.frame(JonesRecords), path = paste0("data/",
                         "JonesRecords", 
                         base::noquote(lubridate::today()),
                         ".csv"))
writeLines(fs::dir_ls("data/"))
```
