---
title: "StorybenchR - tutorials for data storytelling with R"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```


# Welcome to `StorybenchR`

This is a series of tutorials from Data Journalism in R from [Storybench.org](http://www.storybench.org/category/data-journalism-in-r/) using the `tidyverse`. 

```{r tidyverse_logo, echo=FALSE}
crayon::strip_style(tidyverse::tidyverse_logo())
```

## Tutorials Outline

Below is the outline for the *Data Journalism in R* tutorials. More lessons are added as new packages come out or better workflows/methods are discovered. Previous Versions are in the `archive` folder.

***

# Table of Contents

## Title: Getting Started With R in RStudio Notebooks (R markdown) 

### Description: 

Rmarkdown is the foundation for sharing/using code.

###  Orginal post: 

http://www.storybench.org/getting-started-r-rstudio-notebooks/ 

### Github repo: 

https://github.com/mjfrigaard/storybenchR/tree/master/01-r-markdown

## Title: Getting Started with tidyverse in R

Intro to data structures and tidy data format. 

###  Orginal post:

http://www.storybench.org/getting-started-with-tidyverse-in-r/


### Github repo: 

https://github.com/mjfrigaard/storybenchR/tree/master/02.1-tidyr-tidyverse

This has been updated with `pivot_` functions. 

## Title: How to explore and manipulate a dataset from the fivethirtyeight package in R


###  Orginal post:

http://www.storybench.org/how-to-explore-a-dataset-from-the-fivethirtyeight-package-in-r/

### Github repo: 

This was updated to include `ggplot2::qplot()`.

https://github.com/mjfrigaard/storybenchR/tree/master/03.2-tidyr-2-qplot

## Title: How to manipulate data with dplyr in R

###  Orginal post:

http://www.storybench.org/how-to-manipulate-data-with-dplyr-in-r/

### Github repo:

This was broken into two tutorials:

One for manipulating variables:
https://github.com/mjfrigaard/storybenchR/tree/master/05.1-dplyr-vars

And another for observations:
https://github.com/mjfrigaard/storybenchR/tree/master/05.2-dplyr-obs

## Title: Getting started with data visualization in R using ggplot2

### Original post: 

http://www.storybench.org/getting-started-data-visualization-r-using-ggplot2/

### Github repo:

https://github.com/mjfrigaard/storybenchR/tree/master/06.1-ggolot2-adv

## Title: Twitter + R

### Original post: 

http://www.storybench.org/twitter-r/

### Github repo:

This was not put on Github. Contact author at: http://www.storybench.org/author/jonfitz/

## Title: Sentiment analysis of (you guessed it!) Donald Trump’s tweets

### Original post:

http://www.storybench.org/sentiment-analysis-of-you-guessed-it-donald-trumps-tweets/

### Github repo:

This was not put on Github. Contact author at: http://www.storybench.org/author/jonfitz/

## Title: How to merge and clean up multiple CSVs using R

### Original post:

http://www.storybench.org/merge-clean-multiple-csvs-using-r/

### Github repo:

## Title:

### Original post:

### Github repo:



