---
title: "StorybenchR - tutorials for data storytelling with R"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```


# Welcome to `StorybenchR`

This is a series of tutorials from Data Journalism in R from [Storybench.org](http://www.storybench.org/category/data-journalism-in-r/) using the `tidyverse`. 

```{r tidyverse_logo, echo=FALSE}
crayon::strip_style(tidyverse::tidyverse_logo())
```

## Tutorials Outline

Below is the outline for the *Data Journalism in R* tutorials. More lessons are added as new packages come out or better workflows/methods are discovered. Previous Versions are in the `archive` folder.

***

# Table of Contents

## 1) Getting Started With R in RStudio Notebooks (R markdown) 

### Description: 

Rmarkdown is the foundation for sharing/using code.

###  Orginal post: 

http://www.storybench.org/getting-started-r-rstudio-notebooks/ 

### Github repo: 

https://github.com/mjfrigaard/storybenchR/tree/master/01-r-markdown

## 2) Getting Started with tidyverse in R

Intro to data structures and tidy data format. 

###  Orginal post:

http://www.storybench.org/getting-started-with-tidyverse-in-r/


### Github repo: 

https://github.com/mjfrigaard/storybenchR/tree/master/02.1-tidyr-tidyverse

This has been updated with `pivot_` functions. 


