---
title: "Martin's Rmarkdown Notebook Template"
author: "Martin Frigaard"
date: ""
output: 
  html_notebook: 
    highlight: tango
    number_sections: yes
    theme: yeti
    toc: yes
    toc_depth: 6
editor_options: 
    chunk_output_type: inline
    fig_crop: no
    fig_height: 3
    fig_width: 2
    highlight: tango
    toc: yes
    toc_depth: 5
---


```{r setup, include=FALSE}
require(tidyverse)
require(mosaic)
require(magrittr)
knitr::opts_chunk$set(
    echo = TRUE, # show all code
    tidy = FALSE, # cleaner code printing
    size = "small") # smaller code
options(width = 80)
```

This code chunk will create a file name for the document. 

```{r filename}
# filename ----------
file_prefix <- c("001-") # version #
file_exten <- c(".Rmd")
file_title <- tolower(str_replace_all(
    "Martin's Rmarkdown Template", 
    pattern = " ",
    replacement = "_"))
file_name <- paste0(file_prefix, file_title, file_exten)
file_name
```

**PACKAGES:**  
```{r packages}
library(dplyr) # Data wrangling, glimpse() and tbl_df().
library(ggplot2) # Visualise data.
library(lubridate) # Dates and time.
library(readr) # Efficient reading of CSV data.
library(stringr) # String operations.
library(tibble) # Convert row names into a column.
library(tidyr) # Prepare a tidy dataset, gather().
library(magrittr) # Pipes %>%, %T>% and equals(), extract().
library(tidyverse) # all tidyverse packages
library(mosaic) # favstats and other summary functions
library(fs) # file management functions
library(fivethirtyeight)
```

**HEADER:**  

* **File name:** `r file_name`    
* **Created date:** `r Sys.Date()`    
* **R version:** `r R.version.string`   
* `tidyverse` package version: `r packageVersion("tidyverse")` 
* `magrittr` package version: `r packageVersion("magrittr")`  
* `mosaic` package version: `r packageVersion("mosaic")`  
* `dplyr` package version: `r packageVersion("dplyr")`
* `ggplot2` package version: `r packageVersion("ggplot2")`
* `lubridate` package version: `r packageVersion("lubridate")`
* `readr` package version: `r packageVersion("readr")`
* `stringr` package version: `r packageVersion("stringr")`
* `tibble` package version: `r packageVersion("tibble")`
* `tidyr` package version: `r packageVersion("tidyr")`
* `fs` package version: `r packageVersion("fs")`

# Heading 1

Some regular text

**some bold text**

*some italic text*

Tidying Bob Ross data from `fivethirtyeight`...

```{r bob_ross, tidy=TRUE}
library(tidyr)
library(dplyr)
bob_ross <- fivethirtyeight::bob_ross %>% as_tibble()
bob_ross %>% glimpse()
```

Make it tidy and see what `tidy=TRUE` does. 

```{r bob_ross_tidy, tidy=TRUE}
# To convert data frame to tidy data (long) format, run:
library(tidyverse)
library(stringr)
bob_ross_tidy <- bob_ross %>%
  gather(object, present, -c(episode, 
                             season, 
                             episode_num, 
                             title)) %>%
  mutate(present = as.logical(present)) %>%
  arrange(episode, object)
bob_ross_tidy %>% glimpse()
```


## Heading 2

Some regular text

A [link](http://stackoverflow.com/questions/15668870/) to demo below.

```{r grades}
library(tidyr)
library(dplyr)
grades <- tbl_df(read.table(header = TRUE, text = "
   ID   Test Year   Fall Spring Winter
    1   1   2008    15      16      19
    1   1   2009    12      13      27
    1   2   2008    22      22      24
    1   2   2009    10      14      20
    2   1   2008    12      13      25
    2   1   2009    16      14      21
    2   2   2008    13      11      29
    2   2   2009    23      20      26
    3   1   2008    11      12      22
    3   1   2009    13      11      27
    3   2   2008    17      12      23
    3   2   2009    14      9       31"))
grades %>% glimpse()
```

Check out `knitr::kable()` and `tidy=FALSE`. 

```{r grades_gather_mutate_spread_arrage}
knitr::kable(
grades %>%
  gather(Semester, Score, Fall:Winter) %>%
  mutate(Test = paste0("Test", Test)) %>%
  spread(Test, Score) %>%
  arrange(ID, Year, Semester) %>% 
  count(Semester, Year) %>% 
  spread(Year, n))
```


### Heading 3

Some regular text

See what `echo=FALSE` does:

```{r results, echo=FALSE}
library(tidyr)
library(dplyr)
results <- data.frame(
  Ind = paste0("Ind", 1:10),
  Treatment = rep(c("Treat", "Cont"), each = 10),
  value = 1:20
)
results %>% head(5)
```

See what `results='hide'` does: 

```{r results_spread, results='hide'}
results %>% spread(Treatment, value) %>% head(5)
```


#### Heading 4 (about same size as regular text)

Some regular text

##### Heading 5 (smaller than regular text)

Some regular text

###### Heading 6 (smaller than regular text)

Some more regular text

*** 

> Gather example

```{r tidy_example}
set.seed(10)
activities <- data.frame(
  id = sprintf("x1.%02d", 1:10),
  trt = sample(c('cnt', 'tr'), 10, T),
  work.T1 = runif(10),
  play.T1 = runif(10),
  talk.T1 = runif(10),
  work.T2 = runif(10),
  play.T2 = runif(10),
  talk.T2 = runif(10)
)
activities %>% glimpse()
```

Check out the `gather()` %>% `separate()` %>% `arrange()` %>% `spread()`. This is what `tidy=FALSE` looks like. 

```{r gather_separate_arrange_spread}
activities %>%
  gather(key, value, -id, -trt) %>%
  separate(key, into = c("location", "time")) %>%
  arrange(id, trt, time) %>%
  spread(location, value) %>% glimpse(78)
```

**FOOTER:**

```{r session_info, echo=TRUE}
devtools::session_info()  # put this at the end of document
```